AA YMCA current data status
==============

`r opts_chunk$set(tidy = FALSE)`

```{r loadLibraries,echo=F,message=F}
library(data.table)
library(ggplot2)
library(gridExtra)
```
Using: YMCA-DATA/'Original Data - Cleaned.txt'

```{r showData,eval=F}
ydata <- read.delim('../YMCA-DATA/Original Data - Cleaned.txt', header=F, stringsAsFactors=F)
```
```{r loadData,eval=T,echo=F,cache=F}
ydata <- read.delim('/Users/elw/Dropbox/eMacs/projects/aaymca/YMCA-DATA/Original Data - Cleaned.txt', header=F, stringsAsFactors=F)
colnames(ydata) <- c("mem_number_hash","MIN_FAMILY","CHI_DATE","EXP_DATE","MTP_MTY_PERIOD","MTP_MTY_TYP","MTP_PPL_DESCRIP","MST_MEM_STATE")
```

Convert to data.table

```{r convertToDataTable,eval=T}
ydata <- data.table(ydata); str(ydata)
```

<!-- Clean up dates -->

```{r cleanUpDates,eval=T,echo=F,results='hide',cache=F}
	ydata[,CHI_DATE:=as.Date(CHI_DATE, format='%m/%d/%Y %H:%M:%S')]
	ydata[,EXP_DATE:=as.Date(EXP_DATE, format='%m/%d/%y')]
	ydata[,MTP_MTY_PERIOD:=as.numeric(strtrim(ydata$MTP_MTY_PERIOD,4))]
```

<!-- Define useful functions -->

```{r functionChunk,eval=T,echo=F,cache=F}
	
	nthSmallest <- function(vec, n=2){
		sort(vec)[n]
	}

	hasNEntry <- function(vec,n=1){
		return (sum(vec,na.rm=T)==n);
	}

	stateDate <- function(date_vec, state_vec,state){
		sd=c(Inf);
		if(!identical(state_vec[state_vec==state],character(0))){
			sd=date_vec[state_vec[state_vec==state]];
		}
		return(sd);
	}
```
<!-- Skim Data -->

```{r skimData,eval=T,echo=F,message=F,warnings=F,cache=F}
ydata <- suppressWarnings(ydata[,list(secondSmallestCD=nthSmallest(CHI_DATE,2),minCD=min(CHI_DATE), maxCD=max(CHI_DATE),minED=min(EXP_DATE),maxED=max(EXP_DATE),minMTY=min(MTP_MTY_PERIOD,na.rm=T), hasOneNew=hasNEntry(MST_MEM_STATE=="New",1),hasOneExp=hasNEntry(MST_MEM_STATE=="Expired")),by=mem_number_hash]);
```

Meaning of entry labels needs validation/clarification!

Best guesses:
- `mem_number_hash`: Member ID
- `CHI_DATE`: Date of data point entry into system 
	- First entry corresponds to beginning of membership?
- `EXP_DATE`: Date of expiration of membership
	- not always present
- `MTP_MTY_PERIOD`: Tax year of membership(?)
- `MST_MEM_STATE`: Status of membership at time of data point entry

Some summary plots before data cleaning:

```{r sumPlots_1,eval=T,echo=F,results='hide'}
sp1_1=qplot(mem_number_hash,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
sp1_2=qplot(minCD, data=ydata, geom="histogram",xlim=c(as.Date('1985/1/1'),as.Date('2014/1/1')),binwidth=365);
sp1_3=qplot(mem_number_hash,minMTY,data=ydata);
sp1_4=qplot(minMTY,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
```

```{r multiPlots_1,echo=F,dev='pdf'}
	grid.arrange(sp1_1, sp1_2, sp1_3, sp1_4, ncol=2);
```


	Require at least one `CHI_DATE` entry for each member ID.
```{r skimOneChiDate,eval=T,echo=F}
	ydata <- ydata[!is.na(ydata$minCD),];
```
```{r sumPlots_2,eval=T,echo=F,results='hide'}
sp2_1=qplot(mem_number_hash,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
sp2_2=qplot(minCD, data=ydata, geom="histogram",xlim=c(as.Date('1985/1/1'),as.Date('2014/1/1')),binwidth=365);
sp2_3=qplot(mem_number_hash,minMTY,data=ydata);
sp2_4=qplot(minMTY,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
```

```{r multiPlots_2,echo=F,dev='pdf'}
	grid.arrange(sp2_1, sp2_2, sp2_3, sp2_4, ncol=2);
```

	Require at least two `CHI_DATE` entries (ensure follow-up)
```{r skimTwoChiDate,eval=T,echo=F}
	ydata <- ydata[!is.na(ydata$secondSmallestCD),];
```
```{r sumPlots_3,eval=T,echo=F,results='hide'}
sp3_1=qplot(mem_number_hash,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
sp3_2=qplot(minCD, data=ydata, geom="histogram",xlim=c(as.Date('1985/1/1'),as.Date('2014/1/1')),binwidth=365);
sp3_3=qplot(mem_number_hash,minMTY,data=ydata);
sp3_4=qplot(minMTY,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
```

```{r multiPlots_3,echo=F,dev='pdf'}
	grid.arrange(sp3_1, sp3_2, sp3_3, sp3_4, ncol=2);
```


	Remove all Y2K entries (unknown source)
```{r skimY2K,eval=T,echo=F}
	ydata <- ydata[minCD!='2000-01-01',]# d <- d[!is.na(d$MTP_MTY_PERIOD),] # some sort of weirdness here
```
```{r sumPlots_4,eval=T,echo=F,results='hide'}
sp4_1=qplot(mem_number_hash,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
sp4_2=qplot(minCD, data=ydata, geom="histogram",xlim=c(as.Date('1985/1/1'),as.Date('2014/1/1')),binwidth=365);
sp4_3=qplot(mem_number_hash,minMTY,data=ydata);
sp4_4=qplot(minMTY,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
```

```{r multiPlots_4,echo=F,dev='pdf'}
	grid.arrange(sp4_1, sp4_2, sp4_3, sp4_4, ncol=2);
```


	Require *single* `New` and `Exp` entry for each member
```{r skimOneNewExp,eval=T,echo=F}
	ydata <- ydata[hasOneNew==TRUE,];
	ydata <- ydata[hasOneExp==TRUE,];
```
```{r sumPlots_5,eval=T,echo=F,results='hide'}
sp5_1=qplot(mem_number_hash,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
sp5_2=qplot(minCD, data=ydata, geom="histogram",xlim=c(as.Date('1985/1/1'),as.Date('2014/1/1')),binwidth=365);
sp5_3=qplot(mem_number_hash,minMTY,data=ydata);
sp5_4=qplot(minMTY,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
```

```{r multiPlots_5,echo=F,dev='pdf'}
	grid.arrange(sp5_1, sp5_2, sp5_3, sp5_4, ncol=2);
```



	Restrict to `MTP_MTY_PERIOD` > 2004
```{r skimMinMty,eval=T,echo=F}
	cur_name='minMTY'
	set(ydata, which(is.infinite(ydata[[cur_name]])), j = cur_name,value =NA);
	ydata <- ydata[minMTY>=2005,]
```

```{r sumPlots_6,eval=T,echo=F,results='hide'}
	sp6_1=qplot(mem_number_hash,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
	sp6_2=qplot(minCD, data=ydata, geom="histogram",xlim=c(as.Date('1985/1/1'),as.Date('2014/1/1')),binwidth=365);
	sp6_3=qplot(mem_number_hash,minMTY,data=ydata);
	sp6_4=qplot(minMTY,minCD,data=ydata,ylim=c(as.Date('1985/1/1'),as.Date('2014/1/1')));
```

```{r multiPlots_6,echo=F,dev='pdf'}
	grid.arrange(sp6_1, sp6_2, sp6_3, sp6_4, ncol=2);
```
